---
title: "Intro-ggplot2"
author: "Jeff Manson"
date: "12/8/2019"
output: html_document
---

## Attach Packages
```{r}
library(tidyverse)
library(readxl)
library(here)
```

## Read in the data
```{r}

ca_np = read_csv(here("R-workshop-data","ca_np.csv"))
ci_np = read_excel(here("R-workshop-data","ci_np.xlsx"))


```

## MAke a Graph

1. Tell R we're using 'ggplot'
2. Tell R what data we are using and what varibable to plot
3. what type of graph (geom ) we are creating

```{r, echo = FALSE}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()

```

```{r, echo = FALSE}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point()

```


1. create a column graph
2. create a area graph


```{r, echo = FALSE}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_col()


```


```{r, echo = FALSE}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_area()

```



## Intro To customizing 'ggplot' graphs
- color = (update point or line color)
- fill =
- line type
- size
- alhpa - transparentecy


```{r}
ggplot(data = ci_np, aes(x= year, y= visitors)) +
  geom_line(color = "red", 
            linetype = "dashed")

```



```{r}
ggplot(data = ci_np, aes(x= year, y= visitors)) +
  geom_point(color = "blue",
             pch =17,
             size = 4,
             alpha = 0.5)+
  geom_line(color = "red")

```


## Customized elements based on varibale values or outcomes

```{r}
ggplot(data = ci_np, aes(x=year, y= visitors))+
  geom_point(aes(size=visitors, 
                 color = visitors),
             pch = 17,
             alhpa = 0.5)
```

```{r}
ggplot(data = ci_np, aes(x=year, y= visitors))+
  geom_col(aes(fill = visitors), color = "magenta")
           
```

```{r}
ggplot(data = ci_np, aes(x=year, y= visitors))+
  geom_line(size =3,
            color = "red")+
  theme_dark()  
```
## add/update lables and titles

Using 'labs'

```{r}
ggplot(data = ci_np, aes(x=year, y= visitors))+
  geom_point(size =1,
            color = "red")+
  theme_dark() +
  labs(x = "My axis Time",
       y= " annual visitiation",
       title = "channel Islands NP Visitation",
       subtitle = "1963 - 2016")
```

## combine compatible geoms

example pts on lines

```{r}
ggplot(data = ci_np, aes(x=year, y= visitors))+
  geom_line(color = "purple")+
  geom_point(color = "magenta",
             aes(size = year))
```

create a column and line graph
```{r}
ggplot(data = ci_np, aes(x=year, y= visitors))+
  geom_col(fill = "purple")+
  geom_line(color = "magenta")
```


## create line graph with multiple series

```{r}
ggplot(data = ca_np, aes(x = year, y = visitors, color = park_name))+
         geom_line()
```

## facetingg ggplot graphs

- eaxh series gets its own plot panel

```{r}
ggplot(data = ca_np, aes(x = year, y = visitors, color = park_name))+
         geom_line(show.legend = FALSE)+
  theme_minimal()+
  facet_wrap(~ park_name)
```

## Use ggsave to export a plot

```{r}
ggsave("np_graph.png",dpi =180, width = 8, height =7)
```

