---
title: "Filter Joins (like vlookup)"
author: "Jeff Manson"
date: "12/8/2019"
output: html_document
---


```{r, include=FALSE}
library(tidyverse)
library(readxl)
library(here)

```

read in 3 files
- invert_counts_curated.xlsx
- fish_counts_currated.csv
- Kelp_countes_curated.xlsx

```{r}
inverted_counts <- read_excel(here("R-workshop-data","invert_counts_curated.xlsx"))
  
fish_counts <- read_csv(here("R-workshop-data","fish_counts_curated.csv"))
  
kelp_count_abur <- read_excel(here("R-workshop-data","kelp_counts_curated.xlsX"))
```

Select is for columns
filter is for rows

## Filter to match rows based on charater string
only keep observations from fishcounts where common_name = "garibaldi"


```{r}
fish_garibaldi <- fish_counts %>% 
  filter(common_name == "garibaldi")
```
 
## filter to keep or exclude based on numerics
```{r}
fish_over50 <- fish_counts %>% 
  filter(total_count >=50)
```

Create a subset for fish at "mohk"

```{r}
fish_mohk <- fish_counts %>% 
  filter(site == "mohk")
```

Filter to keep rows that match this or that
Keep observation for garibaldi, blacksmith, or balck surfperch

use the in %in% operatior to look for any matches within a vector
```{r}
fish_3sp <- fish_counts %>% 
  filter(common_name %in% c("garibaldi","blacksmith","black surfperch"))
```


Using OR statments
```{r}
fish_3sp_v2 <- fish_counts %>% 
  filter(common_name == "garibaldi" | common_name == "blacksmith"| common_name== "black surfperch")
```

## Filter to return rows that match multiple varibable
I want a subset called urchin_abur_mohk that 
- site is abur oe mohk AND
-specied is "purple urchin"

Some ways to this
- seperate "AND" conditionws withing a single filter line by a comma,
- seperate "and" conditions withing a signle filer by &
- multiple filter statements with %>%

```{r}
urchin_abur_mohk <- inverted_counts %>% 
  filter(site == "abur" | site == "mohk",common_name == "purple urchin")
```

